---
title: "Methodology: Functionals"
date: 2017-07-10
output: md_document
---

[Click here to return to the main README file.](./README.md)

## Introduction

This example imports the CSV files, using a more complex functionals
approach than what we saw in [functionals.md](./functionals.md). It is
more complex, because it uses a custom written function. This
function, `import_data()` imports a CSV file and creates a new column
called file_name to show which file the data came from. If you would
like to see how this function works,
see [here](./R/import_data.R). I'm going to keep this document simple,
and just focus on solving the first extra credit challenge.

The other code is similar to what you've seen before, but is less documented.

```{r code, message=FALSE}

## Init ------------------------------------------------------------------------
## I always put my custom functions in a folder called R. Just a habit.
source("R/import_data.R")
library(dplyr)
library(pander)
if (exists("foo_data")) rm(foo_data)
csv_files <- dir("data/", pattern="csv", full.names = TRUE)

## Imports the CSV data --------------------------------------------------------
foo_data <-  bind_rows(lapply(X=csv_files, FUN=import_data))

```

At this point, we have a data.frame called `foo_data` with `r
nrow(foo_data)` rows of data, just like we did in the other examples. 

## Extra Credit #1

This was the content of the original challenge:

> There are several numeric columns of data. These columns were made by
> sampling a sequence of numbers. But I made an error in one file and
> there is a column of data with a different average in that file than in
> any other file. Can you figure out which file I made the error in?

Unfortuantely, as we shall see, I made a mistake when I tried to make an
intentional mistake and did not set up the data the way I meant to. As 
several of you pointed out to me, the average is basically the same across
each of the data files.

Oops.

```{r ec1, fig.align="center"}

## I only need to look at integer columns, because I am told the error is in
## on of these columns. So, I need to figure out which columns are integers and
## now characters.
column_types <- sapply(foo_data, typeof)
## Only interested in the first three columns.
column_integers <- which(column_types == "integer")

## This lets us keep only the integer columns and the file_name column.
columns_to_select <- c(column_integers,which(names(foo_data) == "file_name"))

file_means <- foo_data %>%
    select(columns_to_select) %>%
    group_by(file_name) %>%
    summarize_all(mean)

pander(data.frame(file_means), caption="These files all have similar means.")

```

The average value for each column in all 20 files is basically the
same. So, we won't find the answer there. What about standard
deviation?

```{r ec1p2, fig.align="center"}

file_sd <- foo_data %>%
    select(columns_to_select) %>%
    group_by(file_name) %>%
    summarize_all(sd)

pander(data.frame(file_sd),
       caption="But they do not the same standard deviation.")

```

To be fair, I messed this up. My goal was to make the AVERAGE in file
four different from the others. But, I made a dumb mistake in the code
I used to produce the data, and the only difference is in the standard
deviation, and even that isn't real big. Sorry! It is a little easier to
see in a box and whiskers plot.

```{r ec1-p3, fig.align="center", fig.width=10}

boxplot(third_col~as.factor(file_name), data=foo_data)

```

If you would like to see the code that produced the data, see this [file](./data/files.R).

## Extra Credit #2

Using basic descriptive statistics, what can you tell me about the
data columns?

```{r ec2}

## OK - I should have called this Extra Credit #1.
## This is much easier than the first challenge.
## And, it works/makes sense!
summary(foo_data)

```
