---
title: "Methodology: Functionals"
date: 2017-07-10
output: md_document
---

[Click here to return to the main README file.](./README.md)

## Introduction

This example imports the CSV files, using a more complex functionals
approach than what we saw in [functionals.md](./functionals.md). It is
more complex, because it uses a custom written function. This
function, `import_data()` imports a CSV file and creates a new column
called file_name to show which file the data came from. If you would
like to see how this function works,
see [here](./R/import_data.R). I'm going to keep this document simple,
and just focus on solving the first extra credit challenge.

The other code is similar to what you've seen before, but is less documented.

```{r code, message=FALSE}

## Init ------------------------------------------------------------------------
## I always put my custom functions in a folder called R. Just a habit.
source("R/import_data.R")
library(dplyr)
library(pander)
if (exists("foo_data")) rm(foo_data)
csv_files <- dir("data/", pattern="csv", full.names = TRUE)

## Imports the CSV data --------------------------------------------------------
foo_data <-  bind_rows(lapply(X=csv_files, FUN=import_data))

```

At this point, we have a data.frame called `foo_data` with `r
nrow(foo_data)` rows of data, just like we did in the other examples. 

## Extra Credit #1

There are several numeric columns of data. These columns were made by
sampling a sequence of numbers. But I made an error in one file and
there is a column of data with a different average in that file than
in any other file. Can you figure out which file I made the error in?

```{r ec1}

## I only need to look at integer columns, because I am told the error is in
## on of these columns. So, I need to figure out which columns are integers and
## now characters.
column_types <- sapply(foo_data, typeof)
## Only interested in the first three columns.
column_integers <- which(column_types == "integer")

## This lets us keep only the integer columns and the file_name column.
columns_to_select <- c(column_integers,which(names(foo_data) == "file_name"))



file_means <- foo_data %>%
    select(columns_to_select) %>%
    group_by(file_name) %>%
    summarize_all(mean)

pander(data.frame(file_means), caption="These files all have similar means.")

file_sd <- foo_data %>%
    select(columns_to_select) %>%
    group_by(file_name) %>%
    summarize_all(sd)

pander(data.frame(file_sd), caption="But not the same standard deviation.")

boxplot(third_col~as.factor(file_name), data=foo_data)

```

To be fair, this was really tricky. I had actually intended the
average to be different, and not just the standard deviation, but I
messed up a little when I created the raw code. Sorry! It is a little
easier to see in a box plot.

```{r ec1-part2}

boxplot(third_col~as.factor(file_name), data=foo_data)

```


## Extra Credit #2

Using basic descriptive statistics, what can you tell me about the
data columns?

```{r ec2}

## OK - I should have called this Extra Credit #1.
## This is much easier than the first challenge.
## Sorry.
summary(foo_data)

```
