---
title: "Methodology: Functionals"
date: 2017-07-10
output: md_document
---

[Click here to return to the main README file.](./README.md)

## Introduction

Like the recursion example, nobody tried to solve the challenge using
a functional methodology. This is the most abstract method we will
discuss today. It is also the best and easiest to type. But first,
let's introduce some new concepts. Most programming languages have
functions, or something similar like a SAS macro. But not all
programming languages treat functions as a first-class citizen. R
does, and this is important. Because a function is just a type of
object, R lets you handle a function like you would any other object,
including passing a function to another function as a parameter.

BOOM! Did you heah just explode? 

Because a functional is a function which accepts a function as an input
parameter.

I know, many of you are probably about to stop reading. So, please,
hold on. The example does make sense.

In the previous examples, we used the function `dir` to find all of
the CSV files in the `data/` folder. To do that, we had to pass `dir`
instructions in the form of parameters. A function is nothing more
than some code which accepts input in a structured manner. It is an
abstraction.

A functional is one step higher in terms of abstraction. A functional
is a function which accepts a function as a parameter. An example
might help. The functional `lapply` will apply an arbitrary function
to every member in a vector. In our previous examples, we always
created a vector called `csv_files`. And we already have a function we
want to use, `read_csv`. Our other examples were simply long-winded
ways of using `read_csv` on every item in `csv_files`. The functional
`lapply` just makes this same idea easier to read/write.


```{r code, message=FALSE}

## Init ------------------------------------------------------------------------
library(readr)   ## Replaces read.csv with read_csv (faster).
if (exists("foo_data")) rm(foo_data)

## Vector of CSV files from data/ ----------------------------------------------
csv_files <- dir("data/", pattern="csv", full.names = TRUE)

## Note: It is not necessary to create a blank data.frame with this method.

## Imports the CSV data --------------------------------------------------------
## This could be done in one step, but it is easier to play with this way.
foo_data <- lapply(X=csv_files, FUN=read_csv)
foo_data <- data.frame(foo_data)

```

This reduces our work to two steps. In practice, I would write this as
a single step `foo_data <- data.frame(lapply(X=csv_files,
FUN=read_csv))`.

The functional `lapply` returns a list. That list has 20 elements in
it, because csv_files has 20 entries. Each element in the list is a
data frame composed of the data from a single file. The function
`data.frame` converts the list to a data frame.

The end result is a single data frame, with all of the data from our
20 files.

Take a look at this code and think about it. This is a high-level idea
and is something even some experienced programmers struggle with. But,
it is an incredibly powerful concept.

**Note:** I know of no way to replicate this in SAS. AFAIK, SAS does
not treat functions as first-class citizens and therefore does not
support this kind of programming technique.


## The Results

```{r results}

print(paste("We imported", nrow(foo_data), "rows of data.", sep=" "))
str(foo_data)

```

[Click here to return to the main README file.](./README.md)
