---
title: "Methodology: Iterative"
date: 2017-07-10
output: md_document
---

[Click here to return to the main README file.](./README.md)

## Introduction

This was _by far_ the most popular solution. An iterative approach is
a form of a loop, which loops over a vector. Examples include `for`
loops and a `while` loop. Although old-school R programmers will cluck
their tongues when contronted with iterative solutions, it is a method
that is familiar to anyone with any extensive programming expensive.

```{r code, message=FALSE}

## Init ------------------------------------------------------------------------
library(readr)   ## Replaces read.csv with read_csv (faster).

## Vector of CSV files from data/ ----------------------------------------------
## NOTE: full.names is your friend! Makes life easier.
csv_files <- dir("data/", pattern="csv", full.names = TRUE)

## Creates an empty data frame -------------------------------------------------
## It is best practive to delete an abject before trying to aned to it.
if (exists("foo_data")) rm(foo_data)
## The code below instantiates a data.frame called foo_data. You do not have to
## define the column types. They will all be NULL and NA, which can be combined,
## logically, with everything.
foo_data <- data.frame()

## Imports the CSV data --------------------------------------------------------
## This will return a lot of feedback from R, telling you the column
## specifications (format) chosen by R for each file.
## The defaults are fine.
## Because we already have a null object (foo_data), we can start rbinding
## immediately.
for(i in seq_along(csv_files)) {
            foo_data <- rbind(foo_data, read_csv(csv_files[i]))
}

```

## The Results

```{r results}

print(paste("We imported", nrow(foo_data), "rows of data.", sep=" "))
str(foo_data)

```

[Click here to return to the main README file.](./README.md)
