---
title: "Methodology: Recursion"
date: 2017-07-10
output: md_document
---

[Click here to return to the main README file.](./README.md)

## Introduction

Nobody tried to solve the challenge using a recursive methodology. A
recursive function is a function which calls itself. Thus, it iterates
over the data like a `for` loop. This methodology is slightly more
complext, but it is better if the import process is complicated.

One big advantage is that it is easier to test a function than it is a
loop.

```{r code, message=FALSE}

## Init ------------------------------------------------------------------------
library(readr)   ## Replaces read.csv with read_csv (faster).
## It is best practice to delete an object before trying to apend to it.
if (exists("foo_data")) rm(foo_data)

## Vector of CSV files from data/ ----------------------------------------------
## NOTE: full.names is your friend! Makes life easier.
csv_files <- dir("data/", pattern="csv", full.names = TRUE)

## Note: It is not necessary to create a blank data.frame with this method.

## Imports the CSV data --------------------------------------------------------

## This is the recursive function which imports the CSV data.
## Lines that start with #' are function documentation.

#' concatenate
#'
#' A simple function which imports all of the CSV files in "data/" and
#' concatenates them together.
#'
#' @param csv_files A vector with all of the CSV files in "data".
#' @param i Our iterator parameter.
#' @return A concatenated data.frame.
#' 
concatenate <- function(csv_files, i) {
    ## Creates a temporary place to put the imported data from file i.
    temp <- read_csv(csv_files[i])
    if (i < length(csv_files)) {
        ## If i < the number of CSV files, we recursively call the function
        ## again, and brind it to the file we just imported above.
        foo_data <- rbind(concatenate(csv_files, i+1), temp)
        return(foo_data)
    } else {
        ## This only happens for the last file.
        return(temp)
    }
} ## End concatenate()

## Uses the function -----------------------------------------
## This function iterates of each file in csv_files and returns a data frame.
foo_data <- concatenate(csv_files, 1)

```

This is a good technique to master. Although this example is a little
arbitrary, there are situations where the complexity introduced by
this approach _are_ well worth the additional time and effort.


## The Results

```{r results, results='as is'}

print(paste("We imported", nrow(foo_data), "rows of data.", sep=" "))
str(foo_data)

```

[Click here to return to the main README file.](./README.md)
